{RandomRationalPoints} package for \emph{Macaulay2}

### 2021-03-17 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt

We thank the reviewers for their helpful suggestions, comments and preliminary codes and examples.

Changes Made to Paper:

a) Section 2.1 (Page 2),

b) 2.2 Page 3,

c) Sentence before Example 2.5 added on Page 4

d) Page 5 Section 5 dimViaBezout added with example 3.1 and paragraph following it.

MAIN CHANGE:

We implement the dimViaBezout to compute the dimension probabilistically. This intersects V(I) with successively higher dimensional random linear spaces until there is an intersection.  For example, if $V(I)$ intersect a random line has a point, then we expect that $V(I)$ contains a hypersurface.  If there was no intersection, this function tries a 2-dimensional linear space, and so on.  This speeds up many computations.
Optional Inputs:

DimensionIntersectionAttempts => ZZ : the number of linear spaces to try before moving to the next dimension
MinimumFieldSize => ZZ: if the ambient field is smaller than this value it will automatically be replaced with an extension. (The user may set the  MinimumFieldSize to ensure that the field being worked over is big enough.  For instance, there are relatively few linear spaces over a field of characteristic 2, and this can cause incorrect results to be provided.)

The reviewer suggested the use of dimDegViaBezout where the use of saturation is needed. We avoid this issue (which may be potentially decelerating) totally by the above successive intersections.

Changes Made to the Source Code:

We have added the MultiplicationStrategy verbatim as provided by the reviewer but have changed the computation of the dimension to be done via dimViaBezout instead of dimDegViaBezout.

New documentation and examples have been provided.

All the strategies now make use of the dimViaBezout to compute the dimension of the given ideal, if the codimension is not already known. In case, that is known, it should be specified along with the call of randomPoints to achieve high speeds.

ProjectionAttempts has been set to 0 in the Default Strategy now.

### 2021-03-18 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt

We thank the reviewers for their helpful suggestions, comments and preliminary codes and examples. We would be very happy to invite the reviewer to be a coauthor to the source code (and also for the paper, if desired).

@@

Optional Inputs:

DimensionIntersectionAttempts => ZZ : the number of linear spaces to try before moving to the next dimension
MinimumFieldSize => ZZ: if the ambient field is smaller than this value it will automatically be replaced with an extension. (The user may set the  MinimumFieldSize to ensure that the field being worked over is big enough.  For instance, there are relatively few linear spaces over a field of characteristic 2, and this can cause incorrect results to be provided.)

The reviewer suggested the use of dimDegViaBezout where the use of saturation is needed. We avoid this issue (which may be potentially decelerating) totally by the above successive intersections.

The MultiplicationTable strategy has also beeen added. (This MultiplicationTable strategy essentially is still a LinearIntersection but it handles the point issue differently. The previous strategy was using decompose to break up the ideal.)

@@

Changes Made to the Source Code:

We have added the MultiplicationStrategy verbatim as provided by the reviewer but have changed the computation of the dimension to be done via dimViaBezout instead of dimDegViaBezout. (This MultiplicationTable strategy essentially is still a LinearIntersection but it handles the point issue differently. The previous strategy was using decompose to break up the ideal.)

### 2021-03-18 FastLinAlg/Paper/RandomRationalPoints.tex
Minor
Comments Added and minor changes to Paper as well

o11 = {{31, -15, 53, 41, 87, 9, -21, -43, -12, 85}}      
    \end{verbatim}
\end{quote}\vspace{-1em}
        }}
    \end{example}
To enable speed ups as indicated in the above discussion, the {\tt ProjectionAttempts} has been set to $0$ in the {\tt Default} strategy so that {\tt GenericProjection} is not opted for unless specified by user.

    \begin{example}
        Finally, we allow our functions to extend our field.
        {{\small\color{blue}
        \begin{quote}
        \begin{verbatim}
i12 : time randomPoints(J, Strategy => LinearIntersection, ExtendField => true)
-- used 0.0612306 seconds

@@

\subsection*{\tt dimViaBezout:} We thank the reviewer for pointing out that in most of the computations, computing the codimension significantly slows down speed. Instead, we compute the dimension probabilistically using {\tt dimViaBezout}. This function takes as input an ideal $I$ in a polynomial ring over a field and intersects $V(I)$ with successively higher dimensional random linear spaces until there is an intersection.  For example, if $V(I)$ intersect a random line has a point, then we expect that $V(I)$ contains a hypersurface.  If there was no intersection, this function tries a $2$-dimensional linear space, and so on.  This speeds up many computations. The function also takes in optional inputs as described below:

@@

As discussed earlier, one can always compute the codimension (say, $c$) of the input ideal using this function, and provide it as an optional input {\tt Codimension => c} whenever calling {\tt randomPoints}. Alternatively, all the strategies for {\tt randomPoints} (see \Cref{strategydetails}) compute the dimension using the above function, which significantly accelerates the computations.

@@

This function takes a polynomial ring as an input and output the coordinate change map,
i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.

@@

\begin{example}	
    We project a curve in 4-space to one 2-space.
    ~~
  {{\small\color{blue}
  \begin{quote}
\begin{verbatim}
i1 : R = ZZ/5[x, y, z, w];

>>>>

o11 = {{31, -15, 53, 41, 87, 9, -21, -43, -12, 85}}      
    \end{verbatim}
\end{quote}\vspace{-2em}
        }}
    \end{example}
To enable speed ups as indicated in the above discussion, the {\tt ProjectionAttempts} has been set to $0$ in the {\tt Default} strategy so that {\tt GenericProjection} is not opted for unless specified by user.
\begin{example}
        Finally, we allow our functions to extend our field.
        {{\small\color{blue}
        \begin{quote}
        \begin{verbatim}
i12 : time randomPoints(J, Strategy => LinearIntersection, ExtendField => true)
-- used 0.0612306 seconds

@@

\subsection*{\tt dimViaBezout:} We thank the reviewer for pointing out that in most of the computations, computing the codimension of the given ideal, significantly slows down speed. Instead, we compute the dimension of $V(I)$ probabilistically using {\tt dimViaBezout}. This function takes as input an ideal $I$ in a polynomial ring over a field and intersects $V(I)$ with successively higher dimensional random linear spaces until there is an intersection.  For example, if $V(I)$ intersect a random line has a point, then we expect that $V(I)$ contains a hypersurface.  If there was no intersection, this function tries a $2$-dimensional linear space, and so on.  This speeds up many computations. The function also takes in optional inputs as described below:

@@

As discussed earlier, one can always compute the codimension (say, $c$) of the input ideal using the output from the above function (or by existing command {\tt codim} beforehand), and provide it as an optional input {\tt Codimension => c} whenever calling {\tt randomPoints}. Alternatively, one can simply rely on any of the strategies for {\tt randomPoints} (see \Cref{strategydetails}) as they compute the dimension of $V(I)$ using {\tt dimViaBezout}, which significantly accelerates the computations.

@@

This function takes a polynomial ring as an input and outputs the coordinate change map,
i.e. given a polynomial ring, this will produce a linear automorphism of the ring.  This function checks whether the map is an isomorphism by computing the Jacobian.

@@

\begin{example}	
    We project a curve in $4$-space to one in $2$-space.
    ~~
  {{\small\color{blue}
  \begin{quote}
\begin{verbatim}
i1 : R = ZZ/5[x, y, z, w];
